\speculake: Hiding Malware in Speculative Execution

### 2018-07-30 paper/related-work.tex
pass on weird machines

\subsection{Weird Machines \& Red-Pills}

%Hardware side effects often leave behind unexpected state transitions 
%that are not, or cannot be, accounted for in modeling environments. Malware
%has historically made use of these bugs to create "weird" models of 
%computation~\cite{weird_machines}. These state transitions can also leak 
%information about the environment in which a program is running, which 
%malware has used to detect honey-pots and emulated 
%environments~\cite{balzarotti2010efficient}.

\textit{Weird machines} are machines composing bugs and features of a system to
create arbitrary computation with interesting
properties~\cite{weird_machines,bratus2011exploit,weird_exploits}. This
encompasses many traditional computer vulnerabilities~\cite{buffer_overflow,
format_string_exploit, shacham2007geometry}. These exploits demonstrate
performing arbitrary computation in places that were not originally intended to
act as computational entities.

Unexpected functionality such as~\cite{bangert2013page} is abundant in modern
processors~\cite{d2015exploiting} as multiple threads share system resources
cooperatively and optimizations are often done across the process
isolation boundary. \speculake takes advantage of cooperative hardware
resources, and constructs a hardware weird machine
using speculative branching and \texttt{Flush+Reload}.

Hardware minutia have also been used to identify and fingerprint execution
environments---in malware this is know as a red-pill~\cite{red-pill} and is
typically used by malware to detect when it is under emulation or a debugging
environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
paleari2009fistful}

>>>>

\subsection{Weird Machines}

\speculake is a \textit{weird machine}---a machine which executes arbitrary
computation using unintended executions paths, often the result of bugs or
attacker exploits~\cite{weird_machines,bratus2011exploit,weird_exploits}. In
particular \speculake takes advantage of speculative execution and a separate
trigger program to execute otherwise dead code.

Previous weird machines have demonstrated many traditional computer
vulnerabilities such as buffer overflows, format string exploits and
return-to-libc attacks without function calls~\cite{buffer_overflow,
format_string_exploit, shacham2007geometry}.

Additionally, unexpected functionality such as instruction-less
computing~\cite{bangert2013page} enables the demonstration of a Turing complete
weird machine by taking advantage of memory translation tables, page-faults, and
hardware exception handling to effectively compute arithmetic and logical
operations without the processor executing any instructions.

These types of machines are abundant in modern processors~\cite{d2015exploiting}
as multiple threads share system resources cooperatively and optimizations are
often done across the process isolation boundary.

Whereas previous weird machines have demonstrated the capability to execute
potentially malicious code in an unexpected manner, \speculake showcases the
ability to execute dead code within the speculative world.

\subsection{Red-Pills}
Hardware minutia have also been used to identify and fingerprint execution
environments---in malware this is know as a red-pill~\cite{red-pill} and is
typically used by malware to detect when it is under emulation or a debugging
environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
paleari2009fistful}
